---
title: "Apriori Algorithm"
author: "Dhrubasattwata Roy Choudhury"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Apriori Algorithm

Apriori is an algorithm for frequent item set mining and association rule learning over relational databases. It proceeds by identifying the frequent individual items in the database and extending them to larger and larger item sets as long as those item sets appear sufficiently often in the database.

```{r}
library('arules')
library('knitr')
library('arulesViz')
```

### Dataset
### Dataset: Importing
We will use the Groceries dataset available within the R environment.
```{r}
data(Groceries)
summary(Groceries)     
```

```{r}
class(Groceries)     
```

```{r}
Groceries@itemInfo[1:20,]    
```

#The following code displays the 1Oth to 20th transactions of the Groceries dataset.
```{r}
apply(Groceries@data[,10:20],2,function(r)  paste(Groceries@itemInfo[r,"labels"],collapse=", "))
```

### BUILD the Model

#Apriori algorithm Implementation below
Finally, run the apriori algorithm on the transactions by specifying minimum values for support and confidence.

### First, get itemsets of length 1
```{r}
itemsets<-apriori(Groceries,parameter=list(minlen=1,maxlen=1,support=0.02,target="frequent itemsets"))
```

```{r}
summary(itemsets)  
```

```{r}
inspect(head(sort(itemsets,by="support"),10))   # lists top 10
```

### Second, get itemsets of length 2
```{r}
itemsets<-apriori(Groceries,parameter=list(minlen=2,maxlen=2,support=0.02,target="frequent itemsets"))
```

```{r}
summary(itemsets)  
```

```{r}
inspect(head(sort(itemsets,by="support"),10))   # lists top 10
```


### Third, get itemsets of length 3
```{r}
itemsets<-apriori(Groceries,parameter=list(minlen=3,maxlen=3,support=0.02,target="frequent itemsets"))
```

```{r}
summary(itemsets)  
```

```{r}
inspect(head(sort(itemsets,by="support"),10))   # lists top 10
```


### Fourth, get itemsets of length 4
```{r}
itemsets<-apriori(Groceries,parameter=list(minlen=4,maxlen=4,support=0.02,target="frequent itemsets"))
```

```{r}
summary(itemsets)  
```

```{r}
inspect(head(sort(itemsets,by="support"),10))   # lists top 10
```

### The Apriori function  () is used to generate rules. A threshold is set lower than 0.001 and minimum confidence threshold is set to 0.6. Below code generates 2,918 rules.
```{r}
rules <- apriori(Groceries,parameter=list(support=0.001,confidence=0.6,target="rules"))
```

```{r}
summary(rules)  
```

```{r}
plot(rules)               # displays scatterplot
```

### Compute the 1/Support(Y) ie slope
```{r}
slope<-sort(round(rules@quality$lift/rules@quality$confidence,2))
```

### Display the number of times each slope appears in dataset
```{r}
unlist(lapply(split(slope,f=slope),length))
```

### Inspect function is used to display the top 10 rules sorted by lift
```{r}
inspect(head(sort(rules,by="lift"),10))
```

```{r}
#Below code fetchces rules with confidence above 0.9
confidentRules<-rules[quality(rules)$confidence>0.9] 
confidentRules       # set of 127 rules
```


```{r}
#Plot a matrix-based visualization of the LHS v RHS of rules. This produces a Matrix-based visualization of LHS and RHS, colored by lift and confidence, 
plot(confidentRules,method="matrix",measure=c("lift","confidence"),control=list(reorder="none"))
```


```{r}
#Visualize the top 5 rules with the highest lift and plot them
highLiftRules<-head(sort(rules,by="lift"),5) 
plot(highLiftRules,method="graph",control=list(type="items"))
```


### ----------------------------------------------------------------------------
### R code to convert a CSV file to a BASKET FORMAT
### ----------------------------------------------------------------------------

```{r}
#read transactions
# df_groceries <- read.csv("Groceries_dataset.csv")
# str(df_groceries)
# df_sorted <- df_groceries[order(df_groceries$Member_number),]

# convert member number to numeric
# df_sorted$Member_number <- as.numeric(df_sorted$Member_number)

#convert item description to categorical format
# df_sorted$itemDescription <- as.factor(df_sorted$itemDescription)
# str(df_sorted)

#convert dataframe to transaction format using ddply; 

# if(sessionInfo()['basePkgs']=="dplyr" | sessionInfo()['otherPkgs']=="dplyr"){
#  detach(package:dplyr, unload=TRUE)
# }

# group all the items that were bought together; by the same customer on the same date
# library(plyr)
# df_itemList <- ddply(df_groceries, c("Member_number","Date"), function(df1)paste(df1$itemDescription,collapse = ","))

#remove member number and date
# df_itemList$Member_number <- NULL
# df_itemList$Date <- NULL

# colnames(df_itemList) <- c("itemList")
```

